---
title: "Analysis of movement patterns of dog walking"
subtitle: "Final Project Patterns and Trends in Environmental Data - Computational Movement Analysis"
author: "Margarita Kessler Rios & Mirjam Nötzli"
format:
  html:
    code-fold: true
---
```{r}
#| warning: false
#| message: false
library(sf)
library(tidyverse)
library(tmap)
library(cowplot)
library(lubridate)
library(dplyr)
library(plotly)
library(tmap)
library(SimilarityMeasures)
library(data.table)
```

#Import data,  handle spatial data and project the coordinates into the new Swiss CRS CH1903+ LV95. Data from two individuals: MK and MM
```{r}
#| warning: false
#| message: false
# Import data
# filter walk transport mode, select only necessary columns, and transform data frame into sf object and transform it into EPSG 2056 coordinate system

#Import Posmo data MK
posmo_MK <- read_delim("data/posmo_MK_edited3.csv", delim = ";") |> 
  filter(transport_mode == "Walk") |> 
  select(user_id, datetime, weekday, lon_x, lat_y, trajectoryID, Type, Shape) |> 
  st_as_sf(coords = c("lon_x","lat_y"), crs = 4326)|>
  st_transform(2056)


# Import Posmo data MN
posmo_MN <- read_delim("data/posmo_MN_edited3.csv", delim = ";") |> 
  filter(transport_mode == "Walk") |> 
  select(user_id, datetime, weekday, lon_x, lat_y, trajectoryID, Type, Shape) |> 
  st_as_sf(coords = c("lon_x","lat_y"), crs = 4326)|>
  st_transform(2056)


str(posmo_MK)
str(posmo_MN)

# import Weather Data with only the relevant columns
precip_data <- read_delim("data/weather_data.csv") |> 
  select(-tre200s0, -sre000z0)

precip_hinwil <- read_delim("data/weather_hinwil.csv")

# filter SHA weather station
precip_SHA <- precip_data |> filter(stn == "SHA") |> rename(datetime = time)
precip_SHA$datetime <- as.POSIXct(precip_SHA$datetime, tz = "UTC") # convert character into POSIXct

# HIW weather station
precip_HIW <- precip_hinwil |> rename(datetime = time) |> rename(rco150z0 = rre150z0)
precip_HIW$datetime <- as.POSIXct(precip_HIW$datetime, tz = "UTC") # convert character into POSIXct
```

#Pre-processing
```{r}
#| warning: false
#| message: false
# store the coordinates into two separate columns
posmo_coordinates <- st_coordinates(posmo_MK)

posmo_MK <- cbind(posmo_MK, posmo_coordinates)

str(posmo_MK)

# the same with MN Data
posmoMN_coordinates <- st_coordinates(posmo_MN)

posmo_MN <- cbind(posmo_MN, posmoMN_coordinates)

```

#calculate longest distance from each trajectory to home for the two data sets
```{r}
#| warning: false
#| message: false
#Create an sf object with the house coordinates
house_MK <- read_delim("data/MK_house.csv") |> 
  st_as_sf(coords = c("lon_x","lat_y"), crs = 4326)|>
  st_transform(2056)

houseMK_coordinates <- st_coordinates(house_MK)

house_MK <- cbind(house_MK, houseMK_coordinates)


#The same with MN 
house_MN <- read_delim("data/MN_house.csv") |> 
  st_as_sf(coords = c("lon_x","lat_y"), crs = 4326)|>
  st_transform(2056)

houseMN_coordinates <- st_coordinates(house_MN)

house_MN <- cbind(house_MN, houseMN_coordinates)

#Calculate the max distance from the house for each trajectory
posmo_MK$distance_home <- as.numeric(st_distance(posmo_MK, house_MK)[,1])
 
posmo_MK <- posmo_MK |>
  group_by(trajectoryID) |>
  mutate(max_distance = max(distance_home))

#the same for MN
posmo_MN$distance_home <- as.numeric(st_distance(posmo_MN, house_MN)[,1])
 
posmo_MN <- posmo_MN |>
  group_by(trajectoryID) |>
  mutate(max_distance = max(distance_home))

```

##Join posmo data with weather data then join the two dataframes
```{r}
#| warning: false
#| message: false
# Create new column with time values rounded to closest 10 minutes, to be able to join with weather data
posmo_MN <- posmo_MN |> mutate(
  Datetime_round = lubridate::round_date(datetime,"10 minutes")
)

posmo_MK <- posmo_MK |> mutate(
  Datetime_round = lubridate::round_date(datetime,"10 minutes")
)
# Join posmo data with weather data
posmo_weather_MN <- left_join(posmo_MN, precip_SHA, by = c("Datetime_round" = "datetime"))

posmo_weather_MK <- left_join(posmo_MK, precip_HIW, by = c("Datetime_round" = "datetime"))

# Delete double timestamps
posmo_weather_MN <- posmo_weather_MN |> 
  distinct(datetime, .keep_all = TRUE)

# Add new column for rain/no rain
posmo_weather_MN$precip[posmo_weather_MN$rco150z0 == 0] <- "no"
posmo_weather_MN$precip[posmo_weather_MN$rco150z0 > 0] <- "yes"

posmo_weather_MK$precip[posmo_weather_MK$rco150z0 == 0] <- "no"
posmo_weather_MK$precip[posmo_weather_MK$rco150z0 > 0] <- "yes"

#Join the two data frames vertically
posmo <- rbind(posmo_weather_MN, posmo_weather_MK)

# Get trajectories with precipitation >50% of the time 
precip50 <- posmo |>
  group_by(trajectoryID) |>
  summarise(percent = length(which(precip == "yes")) / n()) |>
  filter(percent >= 0.5) # trajectories 15, 16, 23, 45, T8, T9, T16, T20

# Create new column "precip_50 for trajectories with >= 50% of the time with rain
posmo <- posmo |> 
  group_by(trajectoryID, precip) |>
  mutate(
     precip_50 = case_when(
         trajectoryID=="15" ~ "yes",
         trajectoryID=="16" ~ "yes",
         trajectoryID=="23" ~ "yes",
         trajectoryID=="45" ~ "yes",
         trajectoryID=="T8" ~ "yes",
         trajectoryID=="T9" ~ "yes",
         trajectoryID=="T16" ~ "yes",
         trajectoryID=="T20" ~ "yes",
     )
)

# assign all other values "no"
posmo$precip_50 <- posmo$precip_50 |> replace_na("no")

# Add a new column wday for Weekday/Weekend
posmo$wday <- ifelse(posmo$weekday %in% c("Sat", "Sun"), "weekend", "weekday") 

#Add a new column start with the time at which the trajectory started and separate the time with format H and M
posmo <- posmo |>
  group_by(trajectoryID) |> 
  mutate(start = min(datetime))

posmo$start <- format(as.POSIXct(posmo$start, format = "%H:%M"))

# Add a new column time_of_day for Morning/Afternoon/Evening
breaks <- hour(hm("3:00", "10:00", "16:00", "21:59"))
labels <- c("Morning", "Afternoon", "Evening")

posmo$time_of_day <- cut(x=hour(posmo$start), breaks = breaks, labels = labels, include.lowest=TRUE)

# Summarise the trajectory shapes
shape_summary <- posmo |>
  group_by(trajectoryID, Shape) |> 
  summarise() |>
  ungroup() |>
  group_by(Shape) |>
  summarise(count = n()) |>
  st_drop_geometry() |>
  mutate(percent = round(count / sum(count) * 100, 2))


# Summarise the walk types (recreational/functional)
Type_summary <- posmo |>
  group_by(trajectoryID, Type) |> 
  summarise() |>
  ungroup() |>
  group_by(Type) |>
  summarise(count = n()) |>
  st_drop_geometry() |>
  mutate(percent = round(count / sum(count) * 100, 2))


# Summarise the walks with precipitation/no Precipitation
precip_summary <- posmo |>
  group_by(trajectoryID, precip_50) |> 
  summarise() |>
  ungroup() |>
  group_by(precip_50) |>
  summarise(count = n()) |>
  st_drop_geometry() |>
  mutate(percent = round(count / sum(count) * 100, 2))

```

#EDA. Plotting all the trajectories with Tmap to identify points that don't belong into the trajectores
```{r}
#| echo: false
#| warning: false
#| message: false
tmap_mode("view")
T1 <- posmo %>%
  filter(trajectoryID == "T1")
T1 |> tm_shape() + tm_dots()

T2 <- posmo %>%
  filter(trajectoryID == "T2")
T2 |> tm_shape() + tm_dots()

T3 <- posmo %>%
  filter(trajectoryID == "T3")
T3 |> tm_shape() + tm_dots()

T4 <- posmo %>%
  filter(trajectoryID == "T4")
T4 |> tm_shape() + tm_dots()

T5 <- posmo %>%
  filter(trajectoryID == "T5")
T5 |> tm_shape() + tm_dots()

T6 <- posmo %>%
  filter(trajectoryID == "T6")
T6 |> tm_shape() + tm_dots()

T7 <- posmo %>%
  filter(trajectoryID == "T7")
T7 |> tm_shape() + tm_dots()

T8 <- posmo %>%
  filter(trajectoryID == "T8")
T8 |> tm_shape() + tm_dots()

T9 <- posmo %>%
  filter(trajectoryID == "T9")
T9 |> tm_shape() + tm_dots()

T10 <- posmo %>%
  filter(trajectoryID == "T10")
T10 |> tm_shape() + tm_dots()

T29 <- posmo %>%
  filter(trajectoryID == "T29")
T29 |> tm_shape() + tm_dots()

T48 <- posmo %>%
  filter(trajectoryID == "T48")
T48 |> tm_shape() + tm_dots()

T39 <- posmo %>%
  filter(trajectoryID == "T39")
T39 |> tm_shape() + tm_dots()



N12 <- posmo %>%
  filter(trajectoryID == "7")
N12|> tm_shape() + tm_dots()

```

#Remove static points from trajectories. posmo-filter contains the trajectories without static points
```{r}
#| warning: false
#| message: false
str(posmo)

#To calculate distance among  Time-stamp 1 and 2 back and forward.
posmo <- posmo |>
  group_by(trajectoryID) |> 
  mutate(
      n_plus1 = sqrt ((lead(X, 1) - X)^2 + (lead(Y, 1)-Y)^2), 
      n_plus2 = sqrt ((lead(X, 2) - X)^2 + (lead(Y, 2)-Y)^2),
      nMinus1 = sqrt((lag(X, 1) - X)^2 + (lag(Y, 1) - Y)^2),
      nMinus2 = sqrt((lag(X, 2) - X)^2 + (lag(Y, 2) - Y)^2)
)

 posmo <- posmo  |>
    rowwise() |>
    mutate(
        stepMean = mean(c(nMinus2, nMinus1, n_plus1, n_plus2))
    ) |>
    ungroup()
 
 ggplot(posmo, aes(stepMean)) +
  geom_histogram(binwidth = 5) +
  geom_vline(xintercept = mean(posmo$stepMean, na.rm = TRUE))
 
which(posmo$stepMean> 100)
which(posmo$stepMean> 50 )
 
#Remove static points by applying threshold,  the threshold is the average step Mean. 
 posmo <- posmo |> 
  mutate(static = stepMean < mean(stepMean, na.rm = TRUE))

posmo_filter <- posmo |>
    filter(!static)


```

# Visualization of all trajectories for both users
```{r}
#| warning: false
#| message: false
tmap_mode("view") 
posmo |> tm_shape() + tm_dots()# both users

posmo |> filter(user_id == "Margarita") |> tm_shape() + tm_dots()# Margarita

posmo |> filter(user_id == "Mirjam") |> tm_shape() + tm_dots()# Mirjam

# Visualization of all functional trajectories of MN
plotMN_functional <- posmo_MN |> 
  filter(Type == "functional") |> 
  ggplot(aes(X, Y, group = trajectoryID)) +
  geom_path(col = "black", alpha = 0.3, size = 0.6) +
  coord_equal(ratio = 1) +
  theme(legend.position = "none") +
  theme_minimal()+
  theme(panel.grid = element_blank(), 
    axis.ticks = element_line(),
    axis.text.y = element_text(angle = 90, hjust = 0.5),
    axis.title.x = element_text(margin = margin(t = 10,r = 0,b=0,l=0)),
    axis.title.y = element_text(margin = margin(t = 0,r = 10,b=0,l=0))) +
  labs(title = "", x = "X", y = "Y")

# Visualization of all functional trajectories of MN
plotMN_recreational <- posmo_MN |> 
  filter(Type == "recreational") |> 
  ggplot(aes(X, Y, group = trajectoryID)) +
  geom_path(col = "black", alpha = 0.3, size = 0.6) +
  coord_equal(ratio = 1) +
  theme(legend.position = "none") +
  theme_minimal()+
  theme(panel.grid = element_blank(), 
    axis.ticks = element_line(),
    axis.text.y = element_text(angle = 90, hjust = 0.5),
    axis.title.x = element_text(margin = margin(t = 10,r = 0,b=0,l=0)),
    axis.title.y = element_text(margin = margin(t = 0,r = 10,b=0,l=0))) +
  labs(title = "", x = "X", y = "Y")

# MK home trajectories (excluding two trajectories in Kanton of Graubünden)
posmoMK_home <- filter(posmo_MK, !trajectoryID %in% c("T2", "T3"))

# Visualization of all functional home trajectories of MK
plotMK_functional <- posmoMK_home |> 
  filter(Type == "functional") |> 
   ggplot(aes(X, Y, group = trajectoryID)) +
  geom_path(col = "black", alpha = 0.3, size = 0.6) +
  coord_equal(ratio = 1) +
  theme(legend.position = "none") +
  theme_minimal()+
  theme(panel.grid = element_blank(), 
    axis.ticks = element_line(),
    axis.text.y = element_text(angle = 90, hjust = 0.5),
    axis.title.x = element_text(margin = margin(t = 10,r = 0,b=0,l=0)),
    axis.title.y = element_text(margin = margin(t = 0,r = 10,b=0,l=0))) +
  labs(title = "", x = "X", y = "Y")

# Visualization of all recreational home trajectories of MK
plotMK_recreational <- posmoMK_home |> 
  filter(Type == "recreational") |> 
   ggplot(aes(X, Y, group = trajectoryID)) +
  geom_path(col = "black", alpha = 0.3, size = 0.6) +
  coord_equal(ratio = 1) +
  theme(legend.position = "none") +
  theme_minimal()+
  theme(panel.grid = element_blank(), 
    axis.ticks = element_line(),
    axis.text.y = element_text(angle = 90, hjust = 0.5),
    axis.title.x = element_text(margin = margin(t = 10,r = 0,b=0,l=0)),
    axis.title.y = element_text(margin = margin(t = 0,r = 10,b=0,l=0))) +
  labs(title = "", x = "X", y = "Y")

# Breaks need to get adjusted to be equal
plot_grid(plotMN_functional, plotMN_recreational, plotMK_functional, plotMK_recreational,
         labels = c("Functional Trajectories", "Recreational Trajectories"),
         hjust = c(-0.3, -0.3),
         vjust = 2,
         ncol = 2, nrow = 2)

```

##Trajectories visualization with and without static points 
```{r}
#| warning: false
#| message: false
## Two from Margarita
K1 <- posmo %>%
  filter(trajectoryID == "T1")

K1_filter <- posmo_filter %>%
  filter(trajectoryID == "T1")
  
K1a <- ggplot(K1, aes(X, Y))+
  geom_path()+
  geom_point(color  = "blue")+
  coord_fixed()

K1b <- ggplot(K1_filter, aes(X, Y))+
  geom_path()+
    geom_point()+
    coord_fixed()

plot_grid(K1a, K1b, 
         labels = c("Raw", "Static points removed"),
          hjust = c(-2,-0.4),
          vjust = 2.5,
          ncol = 2, nrow = 1)

K14 <- posmo %>%
  filter(trajectoryID == "T14")

K14_filter <- posmo_filter %>%
  filter(trajectoryID == "T14")
  
K14a <- ggplot(K14, aes(X, Y))+
  geom_path()+
  geom_point(color  = "blue")+
  coord_fixed()

K14b <- ggplot(K14_filter, aes(X, Y))+
  geom_path()+
    geom_point()+
    coord_fixed()

plot_grid(K14a, K14b, 
         labels = c("Raw", "Static points removed"),
          hjust = c(-2,-0.4),
          vjust = 6,
          ncol = 2, nrow = 1)

## Two from Mirjam
M1 <- posmo %>%
  filter(trajectoryID == "12")


M1_filer <- posmo_filter %>%
  filter(trajectoryID == "12")

M1a <- ggplot(M1, aes(X, Y))+
  geom_path()+
  geom_point(color  = "blue")+
  coord_fixed()

M1b <- ggplot(M1_filer, aes(X, Y))+
  geom_path()+
    geom_point()+
    coord_fixed()

plot_grid(M1a, M1b, 
          labels = c("Raw", "Static points removed"),
          hjust = c(-2,-0.4),
          vjust = 4,
          ncol = 2, nrow = 1)

M26 <- posmo %>%
  filter(trajectoryID == "26")


M26_filter <- posmo_filter %>%
  filter(trajectoryID == "26")
  
M26a <- ggplot(M26, aes(X, Y))+
  geom_path()+
  geom_point(color  = "blue")+
  coord_fixed()

M26b <- ggplot(M26_filter, aes(X, Y))+
  geom_path()+
    geom_point()+
    coord_fixed()

plot_grid(M26a, M26b, 
          labels = c("Raw", "Static points removed"),
          hjust = c(-5.1,-1),
          vjust = 1,
          ncol = 2, nrow = 1)

```

#Timelag. 
How often was the position recorded. Calculated for the trajectories with and without (posmo_filter) static points
```{r}
#| warning: false
#| message: false
posmo <- posmo |> 
    group_by(trajectoryID)|> 
    mutate(timelag = as.numeric(difftime(lead(datetime), datetime, units = "secs")))

posmo_filter <- posmo_filter |> 
    group_by(trajectoryID)|> 
    mutate(timelag = as.numeric(difftime(lead(datetime), datetime, units = "secs")))

 ggplot(posmo, aes(timelag)) +
  geom_histogram(binwidth = 5) +
  geom_vline(xintercept = mean(posmo$timelag, na.rm = TRUE))

which(posmo$timelag> 300)#The threshold was set at 300 (15*20), sometimes we had pauses of up to 20 minutes during the walks that recorded the same location several times
which(posmo$timelag> 500)
which(posmo$timelag> 1000)

 
```

##Visualization of the groups of points that show unreaslistic time lags 
```{r}
#| warning: false
#| message: false
tmap_mode("view")
posmo[18755:18765,] |> tm_shape() + tm_dots()

```

#Steplenght, distance covered and speed calculation
```{r}
#| warning: false
#| message: false
#calculation with the whole trajectories 

posmo <- posmo |>
  group_by(trajectoryID) |> 
  mutate(steplenght = sqrt ((X-lead(X))^2 + (Y-lead(Y))^2))|>
  mutate(speed = mean(steplenght/timelag, na.rm = TRUE))|>
  mutate(duration = (max(datetime)- min(datetime))/60)|>
  mutate(distance = sum(steplenght, na.rm = TRUE))


#calculation with the trajectories without static points
posmo_filter <- posmo_filter |>
  group_by(trajectoryID) |> 
  mutate(steplenght = sqrt ((X-lead(X))^2 + (Y-lead(Y))^2))|>
  mutate(speed = mean(steplenght/timelag, na.rm = TRUE))|>
  mutate(duration = (max(datetime)- min(datetime))/60)|>
  mutate(distance = sum(steplenght, na.rm = TRUE))


ggplot(posmo, aes(steplenght)) +
  geom_histogram(binwidth = 5) +
  geom_vline(xintercept = mean(posmo$steplenght, na.rm = TRUE))

which(posmo$steplenght > 30)#30 meters in 15 seconds its an average
which(posmo$steplenght > 100)

# Distance calculations
# calculate average distance for recreational/functional walks
dist_summary <- posmo |>
  group_by(Type, distance) |> 
  summarise() |> 
  st_drop_geometry()
 
 
mean(dist_summary$distance[1:33]) # average distance of functional walks 1322.05 m
mean(dist_summary$distance[34:103]) # average distance of recreational walks 4472.65 m

# calculate average distance for recreational/functional walks with and without precipitation
precip_dist_summary <- posmo |>
  group_by(Type, precip_50, distance) |> 
  summarise() |> 
  st_drop_geometry()  

mean(precip_dist_summary$distance[1:31]) # average distance of functional walks without rain 1335.13 m
mean(precip_dist_summary$distance[32:33]) # average distance of functional walks with rain 1119.34 m
mean(precip_dist_summary$distance[34:97]) # average distance of recreational walks without rain 4479.78 m
mean(precip_dist_summary$distance[98:103]) # average distance of recreational walks with rain 4396.58 m

# calculate average distance for recreational/functional walks for 3 time of day categories (morning/afternoon/evening)
dist_type_tofday_summary <- posmo |>
  group_by(Type, time_of_day, distance) |> 
  summarise() |> 
  st_drop_geometry()  

mean(dist_type_tofday_summary$distance[1:9]) # average distance of functional morning walks 1387.88 m
mean(dist_type_tofday_summary$distance[10:21]) # average distance of functional afternoon walks 1340.87 m
mean(dist_type_tofday_summary$distance[22:33]) # average distance of functional evening walks 1253.86 m
mean(dist_type_tofday_summary$distance[34:67]) # average distance of recreational morning walks 5226.97 m
mean(dist_type_tofday_summary$distance[68:90]) # average distance of recreational afternoon walks 4063.88 m
mean(dist_type_tofday_summary$distance[91:103]) # average distance of recreational evening walks 3223.01 m

# calculate average distance for morning/afternoon/evening walks
dist_timeofday_summary <- posmo |>
  group_by(time_of_day, distance) |> 
  summarise() |> 
  st_drop_geometry()

mean(dist_timeofday_summary$distance[1:43]) # average distance of morning walks 4423.44 m
mean(dist_timeofday_summary$distance[44:78]) # average distance of afternoon walks 3130.28 m
mean(dist_timeofday_summary$distance[79:103]) # average distance of evening walks 2277.82 m

# calculate average distance for weekday/weekend walks
dist_wday_summary <- posmo |>
  group_by(wday, distance) |> 
  summarise() |> 
  st_drop_geometry()

mean(dist_wday_summary$distance[1:71]) # average distance of weekday walks 3565.24 m
mean(dist_wday_summary$distance[72:103]) # average distance of weekend walks 3236.91 m


# Speed calculations
# calculate average speed for recreational/functional walks
speed_summary <- posmo |>
  group_by(Type, speed) |> 
  summarise() |> 
  st_drop_geometry()
 
mean(speed_summary$speed[1:33]) # average speed of functional walks 1.20 m/s
mean(speed_summary$speed[34:103]) # average speed of recreational walks 1.22 m/s

# calculate average speed for recreational/functional walks with and without precipitation
speed_precip_summary <- posmo |>
  group_by(Type, precip_50, speed) |> 
  summarise() |> 
  st_drop_geometry()  

mean(speed_precip_summary$speed[1:31]) # average speed of functional walks without rain 1.19 m/s
mean(speed_precip_summary$speed[32:33]) # average speed of functional walks with rain 1.31 m/s
mean(speed_precip_summary$speed[34:97]) # average speed of recreational walks without rain 1.22 m/s
mean(speed_precip_summary$speed[98:103]) # average speed of recreational walks with rain 1.30 m/s

# calculate average speed for recreational/functional walks for 3 time of day categories (morning/afternoon/evening)
speed_type_tofday_summary <- posmo |>
  group_by(Type, time_of_day, speed) |> 
  summarise() |> 
  st_drop_geometry()  

mean(speed_type_tofday_summary$speed[1:9]) # average speed of functional morning walks 1.29 m/s
mean(speed_type_tofday_summary$speed[10:21]) # average speed of functional afternoon walks 1.18 m/s
mean(speed_type_tofday_summary$speed[22:33]) # average speed of functional evening walks 1.16 m/s
mean(speed_type_tofday_summary$speed[34:67]) # average speed of recreational morning walks 1.21 m/s
mean(speed_type_tofday_summary$speed[68:90]) # average speed of recreational afternoon walks 1.23 m/s
mean(speed_type_tofday_summary$speed[91:103]) # average speed of recreational evening walks 1.24 m/s

# calculate average speed for morning/afternoon/evening walks
speen_timeofday_summary <- posmo |>
  group_by(time_of_day, speed) |> 
  summarise() |> 
  st_drop_geometry()

mean(speen_timeofday_summary$speed[1:43]) # average speed of morning walks 1.23 m/s
mean(speen_timeofday_summary$speed[44:78]) # average speed of afternoon walks 1.21 m/s
mean(speen_timeofday_summary$speed[79:103]) # average speed of evening walks 1.20 m/s

# calculate average speed for weekday/weekend walks
speed_wday_summary <- posmo |>
  group_by(wday, speed) |> 
  summarise() |> 
  st_drop_geometry()

mean(speed_wday_summary$speed[1:71]) # average speed of weekday walks 1.23 m/s
mean(speed_wday_summary$speed[72:103]) # average speed of weekend walks 1.18 m/s


```

##Visualization of the groups of points that show unreaslistic time lags 
```{r}
#| warning: false
#| message: false
#more than 300 m
tmap_mode("view")
posmo[3090:3095,] |> tm_shape() + tm_dots()

```

#Visualization of parameters
```{r}
#| warning: false
#| message: false
ggplot(posmo, mapping = aes(Type, steplenght))+
  geom_boxplot() +
  labs(x = "Type of dog walking", y = "Steplenght")+
  theme_light()  

ggplot(posmo, mapping = aes(Type, speed))+
  geom_boxplot() +
  labs(x = "Type of dog walking", y = "Speed [m/s]")+
  theme_light()

ggplot(posmo_filter, mapping = aes(Type, speed))+
  geom_boxplot() +
  labs(x = "Type of dog walking", y = "Speed [m/s]")+
  theme_light()#use this one for the report

ggplot(posmo, mapping = aes(Type, distance))+
  geom_boxplot() +
  labs(x = "Type of dog walking", y = "Distance [m]")+
  theme_light()

ggplot(posmo, mapping = aes(wday, distance))+
  geom_boxplot() +
  labs(x = "Day of week", y = "Distance [m]")+
  theme_light()

ggplot(posmo_filter, mapping = aes(wday, speed))+
  geom_boxplot() +
  labs(x = "Day of week", y = "Speed [m/s]")+
  theme_light()

ggplot(posmo, mapping = aes(time_of_day, distance))+
  geom_boxplot() +
  labs(x = "Time of day", y = "Distance [m]")+
  theme_light()

ggplot(posmo_filter, mapping = aes(time_of_day, speed))+
  geom_boxplot() +
  labs(x = "Time of day", y = "Speed [m/s]")+
  theme_light()

ggplot(posmo, mapping = aes(precip_50, distance))+
  geom_boxplot() +
  labs(x = "Precipitation", y = "Distance [m]")+
  theme_light()  

ggplot(posmo, mapping = aes(precip_50, speed))+
  geom_boxplot() +
  labs(x = "Precipitation", y = "Speed [m/s]")+
  theme_light()


# Pie chart of trajectory shapes
plot_ly(shape_summary, labels = ~Shape, values = ~percent, type = 'pie') %>%
  layout(title = "Trajectory Shapes",          
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

# Pie chart of type of walk (recreational/functional)
plot_ly(Type_summary, labels = ~Type, values = ~percent, type = 'pie') %>%
  layout(title = "Type of walk",          
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

# Pie chart of walks in precipitation/no precipitation
plot_ly(precip_summary, labels = ~precip_50, values = ~percent, type = 'pie') %>%
  layout(title = "Precipitation",          
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))


# Scatterplot distance vs speed --> use for report
ggplot(posmo, mapping = aes(distance, speed))+
  geom_point() +
  geom_smooth(method=lm, se=FALSE)+
  labs(x = "Distance [m]", y = "Speed [m/s]")+
  theme_light()     
```

## Plotting more trajectories
```{r}
#| warning: false
#| message: false
T2 <- posmo %>%
  filter(trajectoryID == "T2")

ggplot(T2, aes(X, Y))+
  geom_path()+
  geom_point()+coord_fixed()

T3 <- posmo %>%
  filter(trajectoryID == "T3")

ggplot(T3, aes(X, Y))+
  geom_path()+
  geom_point()+coord_fixed()

T8 <- posmo %>%
  filter(trajectoryID == "T8")

ggplot(T8, aes(X, Y))+
  geom_path()+
    geom_point()+coord_fixed()
```

# Frechet Distance
```{r}
#| warning: false
#| message: false
#randomly select trajectories to compare for MN
x <- sample(1:56, 1)
x
# Functional: 4, 36, 55, 21. Recreational: 6, 39, 49, 53

#randomly select trajectories to compare for MK
x1 <- sample(1:48, 1)
x1
#Functional: T11, T35, T43, T16. Recreational: T40, T20, T27, T25


#Create the matrices for the trajectories to compare 

#Four recreational walks for MN
Traj6MN <- posmo |> 
  ungroup() |>
  st_drop_geometry() |>
  filter(trajectoryID == 6)|>
  mutate(Datetime_int = as.integer(datetime))|>
  select(X, Y, Datetime_int)

Traj6MN <- as.matrix(Traj6MN)

Traj39MN <- posmo |> 
  ungroup() |>
  st_drop_geometry() |>
  filter(trajectoryID == 39)|>
  mutate(Datetime_int = as.integer(datetime))|>
  select(X, Y, Datetime_int)

Traj39MN <- as.matrix(Traj39MN)

Traj49MN <- posmo |> 
  ungroup() |>
  st_drop_geometry() |>
  filter(trajectoryID == 49)|>
  mutate(Datetime_int = as.integer(datetime))|>
  select(X, Y, Datetime_int)

Traj49MN <- as.matrix(Traj49MN)

Traj53MN <- posmo |> 
  ungroup() |>
  st_drop_geometry() |>
  filter(trajectoryID == 53)|>
  mutate(Datetime_int = as.integer(datetime))|>
  select(X, Y, Datetime_int)

Traj53MN <- as.matrix(Traj53MN)
#Four functional walks for MN

Traj4MN <- posmo |> 
  ungroup() |>
  st_drop_geometry() |>
  filter(trajectoryID == 4)|>
  mutate(Datetime_int = as.integer(datetime))|>
  select(X, Y, Datetime_int)

Traj4MN <- as.matrix(Traj4MN)

Traj36MN <- posmo |> 
  ungroup() |>
  st_drop_geometry() |>
  filter(trajectoryID == 36)|>
  mutate(Datetime_int = as.integer(datetime))|>
  select(X, Y, Datetime_int)

Traj36MN <- as.matrix(Traj36MN)

Traj55MN <- posmo |> 
  ungroup() |>
  st_drop_geometry() |>
  filter(trajectoryID == 55)|>
  mutate(Datetime_int = as.integer(datetime))|>
  select(X, Y, Datetime_int)

Traj55MN <- as.matrix(Traj55MN)

Traj21MN <- posmo |> 
  ungroup() |>
  st_drop_geometry() |>
  filter(trajectoryID == 21)|>
  mutate(Datetime_int = as.integer(datetime))|>
  select(X, Y, Datetime_int)

Traj21MN <- as.matrix(Traj21MN) 
  
#Four recreational walks for MK

Traj20MK <- posmo |> 
 ungroup() |>
  st_drop_geometry() |>
  filter(trajectoryID == "T20")|>
  mutate(Datetime_int = as.integer(datetime))|>
  select(X, Y, Datetime_int)

Traj20MK <- as.matrix(Traj20MK)

Traj40MK <- posmo |> 
 ungroup() |>
  st_drop_geometry() |>
  filter(trajectoryID == "T40")|>
  mutate(Datetime_int = as.integer(datetime))|>
  select(X, Y, Datetime_int)

Traj40MK <- as.matrix(Traj40MK)

Traj27MK <- posmo |> 
 ungroup() |>
  st_drop_geometry() |>
  filter(trajectoryID == "T27")|>
  mutate(Datetime_int = as.integer(datetime))|>
  select(X, Y, Datetime_int)

Traj27MK <- as.matrix(Traj27MK)

Traj25MK <- posmo |> 
 ungroup() |>
  st_drop_geometry() |>
  filter(trajectoryID == "T25")|>
  mutate(Datetime_int = as.integer(datetime))|>
  select(X, Y, Datetime_int)

Traj25MK <- as.matrix(Traj25MK)

#Four functional walks for MK

Traj11MK <- posmo |> 
 ungroup() |>
  st_drop_geometry() |>
  filter(trajectoryID == "T11")|>
  mutate(Datetime_int = as.integer(datetime))|>
  select(X, Y, Datetime_int)

Traj11MK <- as.matrix(Traj11MK)

Traj35MK <- posmo |> 
 ungroup() |>
  st_drop_geometry() |>
  filter(trajectoryID == "T35")|>
  mutate(Datetime_int = as.integer(datetime))|>
  select(X, Y, Datetime_int)

Traj35MK <- as.matrix(Traj35MK)

Traj43MK <- posmo |> 
 ungroup() |>
  st_drop_geometry() |>
  filter(trajectoryID == "T43")|>
  mutate(Datetime_int = as.integer(datetime))|>
  select(X, Y, Datetime_int)

Traj43MK <- as.matrix(Traj43MK)

Traj16MK <- posmo |> 
 ungroup() |>
  st_drop_geometry() |>
  filter(trajectoryID == "T16")|>
  mutate(Datetime_int = as.integer(datetime))|>
  select(X, Y, Datetime_int)

Traj16MK <- as.matrix(Traj16MK)

```
#Similarity measures: Calculate Frechet distance for 4 randomly selected functional and 4 recreational walks within the same walker. 
```{r}
# MN Functional: 4, 36, 55, 21. Recreational: 6, 39, 49, 53

A<- Frechet(Traj4MN, Traj36MN, testLeash=-1)# 1291697
B <- Frechet(Traj4MN, Traj55MN)# 1898412
C <- Frechet(Traj4MN, Traj21MN)# 752445
D <- Frechet(Traj36MN, Traj55MN)# 606863 
E <- Frechet(Traj36MN, Traj21MN) #539354.1
Y <- Frechet(Traj55MN, Traj21MN) #1146218 

SM1 <- data.frame (
  Frechet = c(A, B, C, D, E, Y),
  Trajpair = c("T4-T36", "T4-55", "T4-T21", "T36-T55", "T36-T21", "T55-T21"),
  Walker = c ("MN", "MN", "MN", "MN", "MN", "MN"), 
  Type = c("Functional", "Functional", "Functional", "Functional", "Functional", "Functional")
)

G <- Frechet(Traj6MN, Traj39MN) #1355605 
H <- Frechet(Traj6MN, Traj49MN) #1641793
I <- Frechet(Traj6MN, Traj53MN) #1747899
J <- Frechet(Traj39MN, Traj49MN) #286763
K <- Frechet(Traj39MN, Traj53MN) #392869.1
L <- Frechet(Traj49MN, Traj53MN)# 106106.4 


SM2 <- data.frame (
  Frechet = c(G, H, I, J, K, L),
  Trajpair = c("T6-T39", "T6-49", "T6-T53", "T39-T49", "T39-T53", "T49-T53"),
  Walker = c ("MN", "MN", "MN", "MN", "MN", "MN"), 
  Type = c("Recreational", "Recreational", "Recreational", "Recreational", "Recreational", "Recreational" )
)

#MK Functional: T11, T35, T43, T16. Recreational: T40, T20, T27, T25
M <- Frechet(Traj11MK, Traj35MK) #106106.4 
N <- Frechet(Traj11MK, Traj43MK) #1997576
O <- Frechet(Traj11MK, Traj16MK) #409705
P <- Frechet(Traj35MK, Traj43MK) #530645 
Q <- Frechet(Traj35MK, Traj16MK) #1057419 
R <- Frechet(Traj43MK, Traj16MK) # 1587871 

SM3 <- data.frame (
  Frechet = c(M, N, O, P, Q, R),
  Trajpair = c("T11-T35", "T11-43", "T11-T16", "T35-T43", "T35-T16", "T43-T16"),
  Walker = c ("MK", "MK", "MK", "MK", "MK", "MK"), 
  Type = c("Functional", "Functional", "Functional", "Functional", "Functional", "Functional")
)

S <- Frechet(Traj40MK, Traj20MK) #1106042 
Z <- Frechet(Traj40MK, Traj27MK) #786486 
U <- Frechet(Traj40MK, Traj25MK) #870824.5
V <- Frechet(Traj20MK, Traj27MK) #319621.3
W <- Frechet(Traj20MK, Traj25MK) #236536
X <- Frechet(Traj27MK, Traj25MK) #85015.44

#T40, T20, T27, T25
SM4 <- data.frame (
  Frechet = c(M, N, O, P, Q, R),
  Trajpair = c("T40-T20", "T40-27", "T40-T25", "T20-T27", "T20-T25", "T27-T25"),
  Walker = c ("MK", "MK", "MK", "MK", "MK", "MK"), 
  Type = c("Recreational", "Recreational", "Recreational", "Recreational", "Recreational", "Recreational" )
)

Frechet <- rbind(SM1, SM2, SM3, SM4)

ggplot(Frechet, mapping = aes(Type, Frechet))+
  geom_boxplot() +
  labs(x = "Type of dog walking", y = "Frechet Distance")+
  theme_light()  

ggplot(Frechet, aes(Trajpair, Frechet))+
  geom_col(aes(fill = Walker))+
  facet_wrap(~ Type, nrow = 2, scales = "free")+
    theme_light()+
  labs(x = "Trajectory pairs", y = "Frechet Distance")+
  theme(axis.text.x = element_text(angle = 90))


```


